---
title: "New York City flights"
description: |
  An analysis of the `nycflights13` datasets. Mandatory project report in Tools for Analytics (R part).
author:
  - name: [Emil Fisker Jensen]
    affiliation: cand.merc (BI)
  - name: [Mathias Kronborg Wiese]
    affiliation: cand.merc (BI)
  - name: [Mathies Skov Jeppesen]
    affiliation: cand.merc (BI)
  - name: [Mikkel Groth B Christensen]
    affiliation: cand.merc (OSCM)
    
date: "`r Sys.Date()`"
repository_url: [https://github.com/Mikkelgbc/Tools-for-Analytics-R-Part.git]
#preview: https://rstudio.github.io/distill/images/javascript-d3-preview.png
creative_commons: CC BY-NC
output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

```{r, include=FALSE}
if (interactive()) setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # set working dir to current file location
knitr::opts_chunk$set(
  cache = TRUE, autodep = TRUE,  # use this option so compile the document faster (you may comment it out at the final report)
  echo = TRUE, 
  layout="l-page", fig.width = 12
  )
# use xaringan::inf_mr() for instant preview
```

<!-- Some css for the comments (delete) -->
```{css, echo=FALSE}
.comment {
  color: #F1948A;
  border-color: black;
  border-style: dashed;
  border-width: thin;
  padding: 10px;
}
```

## Introduction

We consider the datasets available from the package `nycflights13` that contains information about every flight that departed from New York City in 2013. Let us have a look at the datasets. First, we load the packages need for this report:

```{r, cache=FALSE}
library(tidyverse)
library(nycflights13)
library(skimr)
library(knitr)
library(kableExtra)
library(directlabels)
library(patchwork)
library(rmarkdown)
library(lubridate)
```

The datasets in the `nycflights13` package are:

```{r, fig.align='center', echo=FALSE}
res <- data(package = "nycflights13", verbose = T)
res$results %>% 
  as_tibble() %>% 
  select(Dataset = Item, Description = Title) %>% 
  kable() %>% 
  kable_styling(position = "center")
```

Let us try to do some descriptive analytics on the different datasets.



## Flights

I this section we will focus on the `flights` data set, which lists all domestic flights out of the New York area in 2013. We run `skim` to get an overview:

```{r}
skim(flights)
```

The variables in this dataset are:

* `year, month, day` Date of departure
* `dep_time,arr_time` Actual departure and arrival times.
* `sched_dep_time, sched_arr_time` Scheduled departure and arrival times.
* `dep_delay, arr_delay` delays in minutes
* `hour, minute` Time of scheduled departure
* `carrier` carrier abbreviation
* `tailnum` Tail number of plane.
* `flight` flight number.
* `origin, dest` Origin and Destination
* `air_time` Time spent in air.
* `distance` Distance flown.
* `time_hour` scheduled date and hour of flight.

For further details about the dataset see `?flights` or the [online documentation](https://www.rdocumentation.org/packages/nycflights13/versions/1.0.1/topics/flights).

The skim output indicate that some flights are canceled. We remove these observations from the dataset:

```{r}
dat <- flights %>%
  filter(!is.na(dep_time))
```

### Joining datasets

Let us first try to do some [mutating joins](https://bss-osca.github.io/tfa/sec-transform.html#mutating-joins) and combine variables from multiple tables. In `flights` we have flight information with an abbreviation for carrier (`carrier`), and in `airlines` we have a mapping between abbreviations and full names (`name`). You can use a join to add the carrier names to the flight data:

```{r, warning = FALSE}
dat <- dat %>% 
  left_join(airlines) %>% 
  rename(carrier_name = name) %>% 
  print()
```

Note we here join by the column `carrier` represented in both data frames. That is, the default argument `by = c("carrier" = "carrier")` is used. If we want the full name of origin airport, we need to specify which one we want to join to since each flight has an origin and destination `airport`. Afterwards we do the same for the destination airport. 

```{r}
dat <- dat %>% 
  left_join(airports %>% select(faa, name), 
            by = c("origin" = "faa")) %>% 
  rename(origin_name = name) %>% 
  left_join(airports %>% select(faa, name), 
            by = c("dest" = "faa")) %>% 
  rename(dest_name = name) %>% 
  select(month, carrier_name, origin_name, dest_name, sched_dep_time, dep_delay, arr_delay, distance, tailnum) %>% 
  print()
```

We now have the flights data we need stored in the data frame `dat`. Let us try to answer some questions.





### How many flights leave each New York airport for each carrier? 

We first calculate a summary table:

```{r}
dat %>% 
  count(origin_name, carrier_name, sort = TRUE) %>% 
  paged_table()
```

Let us visualize the numbers. First we facet by airport and use `geom_bar`:


```{r, fig.asp=0.75}
dat %>% 
  ggplot(aes(carrier_name)) +
  geom_bar() + 
  facet_grid(rows = vars(origin_name)) + 
  labs(
    title = "Number of flights",
    x = "Carrier",
    y = "Flights"
  ) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

We can also compare the two categorical variables by using `geom_count`:

```{r}
dat %>%
  ggplot(aes(origin_name, carrier_name)) +
  geom_count() +
  labs(
    title = "Number of flights",
    y = "Carrier",
    x = "Departure airport",
    size = "Flights"
  )
```

Finally, we can use a heatmap by using `geom_tile`. In this case, `geom_tile` doesn't offer a way to calculate counts on it's own, so we use the function `count` in our pipe:

```{r}
dat %>%
  count(origin_name, carrier_name) %>%
  ggplot(aes(origin_name, carrier_name, fill = n)) + 
  geom_tile() + 
  labs(
    title = "Number of flights",
    y = "Carrier",
    x = "Departure airport",
    fill = "Flights"
  ) 
```



### How many carrier flights per month?

Summaries are: 

```{r}
dat %>%
  count(month, carrier_name) %>% 
  paged_table()
```
We will try to visualize the numbers using a line plot with carrier as color aesthetic:

```{r}
dat %>%
  count(month, carrier_name) %>%
  ggplot(mapping = aes(month, n, color = carrier_name)) +
  geom_line() +
  geom_point() +
  geom_dl(aes(label = carrier_name), method = list(dl.trans(x = x + .3), "last.bumpup")) +
  scale_x_continuous(breaks = 1:12, limits = c(1,17)) +
  labs(
    title = "Number of flights",
    y = "Flights",
    x = "Month"
  ) +
  theme(legend.position = "none")
```

### Which carriers/airlines have the worst delays?

Note that delays columns are in minutes. We first convert delays to hours:

```{r}
dat <- dat %>% 
  mutate(across(contains("delay"), ~ .x / 60))
```

Next, we answer the question by looking at different measures.

#### Average delay

Let us first have a look at the average departure delay by airline. The `dplyr` package has two functions that make it easy to do that: the `group_by` and the `summarize` functions. We use the two together and groups the rows of the dataset together based on the `carrier` and then uses `summarise` and the `mean` function to calculate the average delay:

```{r}
dat %>%
  group_by(carrier_name) %>%
  summarise(avg_delay = mean(dep_delay, na.rm = TRUE)) %>%
  arrange(desc(avg_delay)) %>%
  paged_table()
```

Note the `mean` function have a `na.rm` argument which ignores the missing values otherwise the average delays could not be calculated. We can visualize our summary (a continuous-categorical comparison) by piping the table into a column plot:

```{r}
dat %>% 
  group_by(carrier_name) %>% 
  summarise(avg_delay = mean(dep_delay, na.rm = TRUE)) %>%
  ggplot(aes(carrier_name, avg_delay)) + 
  geom_col()
```
To get a better visualization we reorder the categorical x-axis by average delay, use the full names of the airlines (which are rotated) and add some informative labels:

```{r, echo=TRUE}
dat %>% 
  group_by(carrier_name) %>% 
  summarise(avg_delay = mean(dep_delay, na.rm = TRUE)) %>%
  ggplot(aes(reorder(carrier_name, desc(avg_delay)), avg_delay)) + 
  geom_col() +
  labs(
    title = "Average delay for each carrier",
    x = "Carrier",
    y = "Average Delay"
  ) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

To conclude, Frontier (F9) and Express Jet (EV) have the highest average delay. However, using `mean` to summarize a value can be dangerous, because it's sensitive to outliers!





#### Variation

We should *always* ask about the variation in the variables in our data sets, but it's especially important to do so if we're going to use averages to summarize them.

First let us calculate the standard deviation for each carrier:

```{r}
dat %>% 
  group_by(carrier_name) %>% 
  summarise(sd_delay = sd(dep_delay)) %>% 
  paged_table()
```

What is the distribution of departure delays by airline? Visualized as a density distribution using carrier as fill aesthetic:

```{r warning=FALSE}
dat %>%
  ggplot(aes(dep_delay, fill = carrier_name)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Distribution of departure delays by carrier",
    x = "Delay (hours)",
    y = "Density",
    fill = "Carrier"
  )
```

We can see that there is a small number of HUGE outliers which makes using `mean` possibly very misleading. 


Lets us try to make a plot of the empirical cumulative distributions for each carrier using carrier as color aesthetic and a zoom of at most 3 hours delay:

```{r}
dat %>%
  ggplot() +
  stat_ecdf(aes(x = dep_delay, color = carrier_name), alpha = 0.75) +
  coord_cartesian(xlim = c(0,3)) +
  labs(
    title = "Distribution of departure delays by carrier",
    x = "Delay (hours)",
    y = "Density",
    color = "Carrier"
  )
```
Note, the higher upper-left the distribution is, the better. That is, a carrier dominates other carriers if the line is above the other carriers. Comparing this to the standard deviations, we see that the standard deviations is not a good measure for delays.

Variation in data like these where the outliers are very sparse is hard to visualize using density plots. We may also use a boxplot:

```{r}
dat %>% 
  ggplot(aes(dep_delay, fill = carrier_name)) +
  geom_boxplot() +
  coord_cartesian(xlim = c(0,3)) +
  theme(axis.text.y = element_blank()) +
  labs(
    title = "Boxplot of departure delays by carrier",
    x = "Delay (hours)",
    fill = "Carrier"
    )
```

We can see that most flights have a median around zero. However, some carriers have larger delays compared to others. Is the variation in departure delay different given departure airport? We use departure airport as color aesthetic:

```{r}
dat %>% 
  group_by(origin_name) %>% 
  summarise(avg_delay = mean(dep_delay), na.rm = TRUE) %>% 
  ggplot(aes(reorder(origin_name, avg_delay), avg_delay, fill = origin_name)) +
  geom_col() +
  labs(
    title = "Departure delay by Airport",
    x = "Airport",
    y = "Average delay (hours)",
    fill = "Airport"
    )
```

We can see that La Guardia has the lowest average departure delay, while Newark has the highest average departure delay.




#### Median

The boxplot shows median values in the center. What would happen if we used `median` instead of average delay time and make a column plot? 

```{r}
dat %>%
  group_by(carrier_name) %>%
  summarise(median_delay = median(dep_delay, na.rm = TRUE)) %>%
  ggplot(aes(reorder(carrier_name, median_delay), median_delay)) +
  geom_col() +
  labs(
    title = "Median departure delay for each carrier",
    x = "Carrier",
    y = "Median delay (hours)"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
That tells a bit of a different story!  The median delay for most airlines are negative, meaning that they depart before scheduled. Only AirTran Airways, Frontier Airlines and Southwest Airlines have positive median delays.



#### Delays of more than an hour

How many flights were **really** delayed and how does that break down by airline carrier?  Being delayed more than an hour really sucks, so let's use that as our cutoff:

```{r}
dat %>% 
  filter(dep_delay > 1) %>% 
  count()
```

That's a lot of flights!  We can use the `dplyr` function named `count` to give us a summary of the number of rows of a that correspond to each carrier:

```{r}
dat %>% 
  filter(dep_delay > 1) %>% 
  count(carrier_name, sort = TRUE)
```

Note that `count` has created a column named `n` which contains the counts and we ask it to sort that column for us.

We can visualize it with a column plot:

```{r}
dat %>% 
  filter(dep_delay > 1) %>% 
  count(carrier_name) %>% 
  ggplot(aes(reorder(carrier_name, -n), n)) +
  geom_col() +
  labs(
    title = "Number of flights delayed more than 1 hour for each carrier",
    x = "Carrier",
    y = "Fligths"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

We see that ExpressJet Airlines, JetBlue Airways and United Air Lines has the highest number of flights delayed more than 1 hour, while Alaska Airlines, Hawaiian Airlines and SkyWest Airlines have the lowest number of flights delayed more than 1 hour.



### What is the relationship between departure delay and arrival delay?

We plot the delays against each other as points. 

```{r}
dat %>% 
  ggplot(aes(dep_delay, arr_delay)) +
  geom_point(alpha = 0.1) +
  labs(
    title = "Relation between Departure delay and Arrival Delay",
    x = "Departure delay",
    y = "Arrival delay"
  )
```

The large mass of points near (0, 0) can cause some confusion since it is hard to tell the true number of points that are plotted. This is the result of a phenomenon called overplotting. As one may guess, this corresponds to points being plotted on top of each other over and over again. When overplotting occurs, it is difficult to know the number of points being plotted. We adjust the transparency of the points by setting `alpha = 0.1`

We see that there are a linear relation between Departure Delay and Arrival Delay, with most flights being delay between 0 and 5 hours, and then some huge outliers of between 10 and 22 hours


### Are flight delays worse at different New York airports? 

If you're flying out of New York you might want to know which airport has the worst delays on average. We first calculate median and average delays:

```{r}
dat %>%
  group_by(origin_name) %>%
  summarize(median_delay = median(dep_delay),
            avg_delay = mean(dep_delay))
```

As we can see La Guardia seems to have the smallest delays. However, the difference is small. Lets us try to make a plot of the empirical cumulative distributions (ECD) for each airport using airport as color aesthetic and a zoom of at most 2 hours:

```{r}
dat %>%
  ggplot() +
  stat_ecdf(aes(x = dep_delay, color = origin_name), alpha = 0.75) +
  coord_cartesian(xlim = c(-0.2,2)) +
  labs(
    title = "Distribution of departure delays by airport",
    x = "Delay (hours)",
    y = "Density",
    color = "Carrier"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

The median values can be found at y = 0.5. Note that La Gaardia is above the other lines indicating that it has the smallest delays no matter what fractile we consider. Another way to visialize this covariation in a categorical (airport) and a continuous (delay) variable is with a boxplot. We use a little scaling to get a better picture of the average delay and zoom so the y variable is between at most half an hour.

```{r}
dat %>% 
  ggplot(aes(origin_name, dep_delay)) +
  geom_boxplot() +
  coord_cartesian(ylim = c(-0.5,0.5)) +
  labs(
    title = "Boxplot of departure delays by airport",
    x = "Airport",
    y = "Delay (hours)"
    )
```


### Are carrier flight delays different at New York airports? 

We first calculate median and average delays:

```{r}
dat %>%
  group_by(carrier_name, origin_name) %>%
  summarise(median_delay = median(dep_delay),
            avg_delay = mean(dep_delay))
```

There are some differences. Let us try to do a heat map of the average delays:

```{r}
dat %>%
  group_by(carrier_name, origin_name) %>%
  summarize(avg_delay = mean(dep_delay)) %>% 
  ggplot(aes(origin_name, carrier_name, fill = avg_delay)) +
  geom_tile() +
  scale_fill_continuous(low = "#31a354", high = "#e5f5e0") +
  labs(
    title = "Average departure delays",
    x = "Departure airport",
    y = "Carrier",
    fill = "Avg. delay (min)"
  )
```
For each carrier this give a good insight into the differences at each airport. Another way to visualize the covariation is with a boxplot. We use a little scaling to get a better picture of the delay and zoom so the delay is a most half an hour.

```{r}
dat %>%
  ggplot(aes(carrier_name, dep_delay, fill = origin_name)) +
  geom_boxplot() +
  coord_cartesian(ylim = c(-0.5,0.5)) +
  labs(
    title = "Average departure delays",
    x = "Airport",
    y = "Average delay",
    fill = "Airport"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

We may also try to plot the empirical cumulative distributions for each carrier (facet) using airport as color aesthetic and a zoom of the delay is at most 1 hour:

```{r}
dat %>%
  ggplot() +
  stat_ecdf(aes(dep_delay, color = origin_name), alpha = 0.75) +
  coord_cartesian(xlim = c(0,1)) +
  facet_wrap(vars(carrier_name)) +
  labs(
    title = "Distribution of departure delays for carriers by airport",
    x = "Delay (hours)",
    y = "Density",
    color = "Airport"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```



### Does departure time affect flight delays? 

First, note that the `sched_dep_time` is a number in the format HHMM. We convert it into a hour:minutes data type and afterwards to hours since midnight:

```{r}
dat <- dat %>% 
  mutate(sched_dep_time = hm(str_replace(sched_dep_time, "^(.*)(..)$", "\\1:\\2"))) %>% 
  mutate(sched_dep_time = as.numeric(sched_dep_time)/60/60)
```


To explore covariation in two continuous (quantitative) variables, we can use a scatter plot:

```{r}
dat %>% 
  ggplot(aes(sched_dep_time, dep_delay)) +
  geom_point() +
  labs(
    title = "Covariance between departure time and delay",
    x = "Departure time",
    y = "Delay"
  )

```

There seems to be more outliers in the early hours from 6 to 11, and again in late afternoon from 16 to 21



### Does travel distance affect departure and arrival delay?

We use the patchwork package to plot distance against the two delays. Moreover we also add a smoothed line using `geom_smooth`:

```{r}
p1 <- dat %>%
  ggplot(aes(distance, dep_delay)) +
  geom_smooth() +
  labs(
    title = "Distance and departure delay",
    x = "Distance",
    y = "Delay"
  )

p2 <- dat %>%
  ggplot(aes(distance, arr_delay)) +
  geom_smooth() +
  labs(
    title = "Distance and arrival delay",
    x = "Distance",
    y = "Delay"
  )

p1 + p2
```

Both in terms of departure and arrival, there seems to be more delay on shorter flights from 0 to 1100 km. Departure delay seems to flatten afterwards, while arrival delay continues to decrease as the flight distance increases. The variability in delay is however increased with distance, as indicated by the gray area (confidence interval)




## Planes

Let us do a mutation join so we have a bit more information about each airplane:

```{r}
dat <- dat %>% 
  left_join(planes %>% 
              select(tailnum, plane_manufacturer = manufacturer, plane_model = model))
```

### What is the monthly usage of all the aircrafts? 

This could be useful for some kind of maintenance activity that needs to be done after x number of trips. The summary table is (based on `tailnum`):

```{r}
dat %>% 
  count(month, tailnum, sort = TRUE)
```

As an example, consider the plane N355NB:

```{r}
dat1 <- dat %>% 
  filter(tailnum=="N355NB") 
```

The specifications are:

```{r}
filter(planes, tailnum=="N355NB")
```

We see that it is an Airbus 319 with 145 seats. The plane flew `r nrow(dat1)` flights in 2013 with a total distance of `r sum(dat1$distance)`.   

Let us have a look at the destinations:

```{r}
dat1 %>%
  count(dest_name) %>% 
  ggplot(aes(n, dest_name)) +
  geom_col() +
  labs(
    title = "Number of flights to each destination in 2013",
    x = "Number of fligths",
    y = "Destination"
  )

dat1 %>%
  ggplot(aes(distance, dest_name)) +
  geom_col() +
  labs(
    title = "Distance travelled to each destination in 2013",
    x = "Total distance",
    y = "Destination"
  )
```


## Weather

I this section we will focus on the `weather` data set, which lists hourly meterological data for LGA, JFK and EWR. We run `skim` to get an overview:

```{r}
skim(weather)
```

For further details see `View(weather)` or read the associated help file by running `?weather` to bring up the help file.

Observe that there is a variable called temp of hourly temperature recordings in Fahrenheit at weather stations near all three major airports in New York City: Newark (origin code EWR), John F. Kennedy International (JFK), and LaGuardia (LGA). Let us transform the temperature to celsius:

```{r}
dat_w <- weather %>% 
  left_join(airports %>% select(faa, name), 
            by = c("origin" = "faa")) %>% 
  rename(origin_name = name) %>% 
  mutate(temp = (temp - 32) * (5/9) ) %>% 
  select(origin_name, time_hour, month, temp)
```

### How are the temperature fluctutating over the year?

We start by plotting temperature over the year with airport/origin as color aesthetic. We also add a smoothing line:

```{r}
dat_w %>% 
  ggplot(aes(time_hour, temp, color = origin_name)) +
  geom_smooth(alpha = 0.2) +
  labs(
    title = "Temperature over the year",
    x = "Time",
    y = "Temperature",
    color = "Airport"
  )
```

Note that we have used the `alpha` aesthetic to make the lines more transparent. We see that all three airport follow the same trend in temperature. However, John F Kennedy Intl does have a slightly lower temperature from april to october compared the others.

### Are the temperatures different in the airports? 

Let us start by plotting the density for each airport:

```{r}
dat_w %>% 
  ggplot(aes(temp, color = origin_name)) +
  geom_density() +
  labs(
    title = "Temperatures over the year at different airports?",
    x = "Temperature",
    y = "Density",
    color = "Airport"
  )
```

Note the mean temperature is more or less the same (vertical lines). There is a bit fluctuations on Newark compared to for instance JFK airport (lowest spread).

A closer look can be done by faceting by month:

```{r}
dat_w %>% 
  ggplot(aes(temp, color = origin_name)) +
  geom_density() +
  facet_wrap(vars(month)) +
  labs(
    title = "Temperatures over the year at different airports?",
    x = "Temperature",
    y = "Density",
    color = "Airport"
  )
```

Finally, let us consider a boxplot of temperature for each month:

```{r}
dat_w %>%
  ggplot(aes(origin_name, temp)) +
  geom_boxplot() +
  facet_grid(cols = vars(month)) +
  labs(
    title = "Boxplot of temperature over the year by airport",
    x = "Airport",
    y = "Temperature"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

We have made a boxplot for each month and can easily see potential outliers for each month. There are many outliers in march, may and october.

What does the dot at the bottom of the plot for May correspond to?  Explain what might have occurred in May to produce this point.

The dot in the boxplot of may is an outlier of -10 degrees in a month where the average temperature is around 20 degrees. This could be explained by a sensor failure, resulting in a false observation.


## Any insights on canceled flights?

The canceled flights are:

```{r}
dat_c <- flights %>%
  filter(is.na(dep_time)) %>% 
  print()
```

Let us do some analysis.

We transform data to add more information to the canceled fligths:

```{r}
dat_o <- dat_c %>% 
  left_join(airlines) %>% 
  rename(carrier_name = name) %>% 
  left_join(airports %>% select(faa, name), 
            by = c("origin" = "faa")) %>% 
  rename(origin_name = name) %>% 
  left_join(airports %>% select(faa, name), 
            by = c("dest" = "faa")) %>% 
  rename(dest_name = name) %>% 
  select(month, carrier_name, origin_name, dest_name) %>% 
  print()
```


### Which carrier has the most cancelations?

```{r}
dat_o %>% 
  group_by(carrier_name) %>% 
  count(carrier_name, sort = TRUE) %>% 
  ggplot(aes(reorder(carrier_name, n), n))+
  geom_col() +
  coord_flip() +
  labs(
    title = "Number of canceled fligths pr. carrier",
    x = "Carrier",
    y = "Number of cancelled flights"
  )
```


### Which month has the most cancelations?

```{r}
dat_o %>% 
  ggplot() +
  geom_bar(aes(month)) +
  scale_x_continuous(breaks = 1:12)+
  labs(
    title = "Cancelations pr. month",
    x = "Month",
    y = "Cancelations"
  )
```


### Which Departure Airport has the most cancelations?

```{r}
dat_o %>% 
  ggplot() +
  geom_bar(aes(origin_name)) +
  labs(
    title = "Cancelations pr. departure airport",
    x = "Departure airport",
    y = "Cancelations"
  )
```

## Other insights?

### Most popular Aircraft Manufacturer

Transform data to merge manufacture variants

```{r}
dat2 <- dat
dat2 [which(dat1 $plane_manufacturer == "AIRBUS INDUSTRIE"),] $plane_manufacturer <- "AIRBUS"
dat2 [which(dat1 $plane_manufacturer == "CANADAIR LTD"),] $plane_manufacturer <- "CANADAIR"
dat2 [which(dat1 $plane_manufacturer %in% c("MCDONNELL DOUGLAS AIRCRAFT CO", "MCDONNELL DOUGLAS CORPORATION")),] $plane_manufacturer <- "MCDONNELL DOUGLAS"
```

```{r}
dat2 %>% 
  group_by(plane_manufacturer) %>% 
  distinct(tailnum) %>% 
  count(plane_manufacturer, sort = TRUE) %>% 
  filter(n > 10) %>% 
  ggplot() +
  geom_col(aes(reorder(plane_manufacturer, n), n)) +
  coord_flip() +
  labs(
    title = "Most popular aircraft manufacturer (> 10)",
    x = "Aircraft Manufacturer",
    y = "Number of Aircrafts"
  )
```

### Most popular plane model for the three biggest manufacturers (Boeing, Airbus and Bombardier)?

```{r}
dat2 %>% 
  group_by(plane_manufacturer) %>%
  distinct(tailnum, .keep_all = TRUE) %>% 
  filter(plane_manufacturer %in% c("BOEING", "AIRBUS", "BOMBARDIER INC")) %>%
  count(plane_model) %>% 
  slice_max(n) %>% 
  ggplot(aes(reorder(plane_model, n), n, fill = plane_manufacturer)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Most popular aircraft model for the three biggest manufactures",
    x = " Aircraft Model",
    y = "Number of aircraft models",
    fill = "Aircraft Manufacturer"
  )
```
### What is the average distance pr. airline?

```{r}
dat %>% 
  group_by(carrier_name) %>% 
  summarise(avg_distance = mean(distance)) %>% 
  ggplot() +
  geom_col(aes(avg_distance, reorder(carrier_name, avg_distance))) +
  labs(
    title = "Average distance pr. airline",
    x = "Average Distance",
    y = "Airline"
  )
```




## Colophon 
<!-- Always keep this section -->

This report has been created inside [RStudio](http://www.rstudio.com/ide/) using [R Markdown](https://rmarkdown.rstudio.com/) and the [distill](https://rstudio.github.io/distill/) format. 

The report was built using:

```{r message = FALSE, warning = FALSE, echo = FALSE}
session <- devtools::session_info()
session$platform
```

Along with these packages:

```{r message = FALSE, warning = FALSE, echo = FALSE}
session$packages %>% 
  select(package, loadedversion, date, source) %>% 
  DT::datatable(rownames = FALSE,
                class = 'cell-border stripe',
                filter = list(position = 'top'),
                options = list(pageLength = 5, 
                           autoWidth = FALSE,
                           bInfo = FALSE,
                           paging = TRUE))
```
